Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-11-12 paper/main.tex
resolved conflict
work on the soundness theorem

Finally, still following previous work, we define the a continuation based
soundness of triples and prove the soundness of all rules in Coq.  The
full version of this paper \cite{anon_extended}, gives more explanations
on the rules and provides a thorough overview of our soundness proof.

@@

We use the
logic described in the previous paragraph as the target of a translation
function for the automatic derivation system.  This reveals two orthogonal
aspects of our proof that relies, on one side, on amortized reasoning
(as encoded in the quantitative logic), and on the other side, on
combinatorial properties of the linear potential functions that we
are using (as encoded in the automatic analysis).

We define a translation
function $\tr$, such that if a judgement $J$ in the automatic analysis
is derivable, $\tr(J)$ is derivable in the quantitative logic.  By using
$\tr$ to translate derivations of the automatic analysis to derivations
in the quantitative logic we can automatically obtain a certified resource
bound for the analyzed program.

The translation of an assertion $(\Gamma, Q)$ in the automatic
analysis is defined by
$$
  \tr(\Gamma, Q) := \lambda\state .\, (\Gamma(\state)) + \Phi_Q(\state),
$$
where we write $\Phi_Q$ for the unique linear potential function
defined by the quantitative annotation $Q$.  We also need to
translate the assumptions in function contexts of the automatic
analysis.  We define $\tr(\Gamma_f, Q_f, \Gamma'_f, Q'_f) := \forall z \, \vec v \, v . $
%
$$
(
  \lambda \vec v \, \state .\, \tr(\Gamma_f, Q_f)(\state[\Vargs {\mapsto} \vec v]),
  \lambda v \, \state .\, \tr(\Gamma'_f, Q'_f)(\state[\Vret {\mapsto} v])
  )
$$
%
These definitions let us translate the judgement $J = B; R; P \vdash
S \dashv P'$ in the context $\Delta$ by $\tr(J) :=$
$$
  (\lambda f .\, \tr(\Delta(f))); \tr(B); \tr(R) \vdash
    \htriple{\tr(P)}{S}{\tr(P')}.
$$
The soundness of the automatic analysis can now be stated
formally with the following theorem.
%
\begin{theorem}[Soundness of the automatic analysis]
  If $J$ is a judgement derived with the rules of
  \pref{fig:auto}, then $\tr(J)$ is a quantitative
  Hoare triple derivable with the rules of
  \pref{fig:logic}.
\end{theorem}
%
\noindent
The proof of this theorem is constructive and basically maps each rule
of the automatic analysis directly to its counterpart in the
quantitative logic.  The most tricky parts are the translations of the
rules for increments and decrements and the rule {\sc Q:Weak} for
weakening because we have to show that the preconditions of the rules
{\sc L:Weak} or {\sc L:Update}, respectively, are met.  \iffull{ For instance
we prove the following lemma to show the soundness of the translation
of {\sc Q:Weak}.

>>>>

Finally, we define a continuation based
soundness of triples and prove the soundness of all rules in Coq.  The
full version of this paper \cite{anon_extended}, gives more explanations
on the rules and provides a thorough overview of our soundness proof.

@@

We use the
logic from previous paragraph as the target of a translation
function for the automatic derivation system.  This reveals two orthogonal
aspects of the proof: on one side, it relies on amortized reasoning
(the quantitative logic rules), and on the other side, it
uses combinatorial properties of our linear potential functions
(the automatic analysis rules).

Technically, we define a translation
function $\tr$ such that if a judgement $J$ in the automatic analysis
is derivable, $\tr(J)$ is derivable in the quantitative logic.  By using
$\tr$ to translate derivations of the automatic analysis to derivations
in the quantitative logic we can directly obtain a certified resource
bound for the analyzed program.

The translation of an assertion $(\Gamma, Q)$ in the automatic
analysis is defined by
$$
  \tr(\Gamma, Q) := \lambda\state .\, (\Gamma(\state)) + \Phi_Q(\state),
$$
where we write $\Phi_Q$ for the unique linear potential function
defined by the quantitative annotation $Q$.  The logical context $\Gamma$
is implicitely lifted to a quantitative assertion by mapping a state
$\state$ to 0 if $\Gamma(\state)$ holds and to $\infty$ otherwise.
\iffull{We also need to
translate the assumptions in function contexts of the automatic
analysis.  We define $\tr(\Gamma_f, Q_f, \Gamma'_f, Q'_f) :=
\forall z \, \vec v \, v . $
%
$$
(
  \lambda \vec v \, \state .\, \tr(\Gamma_f, Q_f)(\state[\Vargs {\mapsto} \vec v]),
  \lambda v \, \state .\, \tr(\Gamma'_f, Q'_f)(\state[\Vret {\mapsto} v])
  )
$$}{}
%
\iffull{These definitions let}{This lets} us translate the judgement
$J := B; R; P \vdash S \dashv P'$\iffull{ in the context $\Delta$ by $\tr(J) :=$}{ by}
\iffull{$$
  (\lambda f .\, \tr(\Delta(f))); \tr(B); \tr(R) \vdash
    \htriple{\tr(P)}{S}{\tr(P')}.
$$}{%
$$
  \tr(J) := \tr(B); \tr(R) \vdash \htriple{\tr(P)} S {\tr(P')}.
$$}
The soundness of the automatic analysis can now be stated
formally with the following theorem.
%
\begin{theorem}[Soundness of the automatic analysis]
  If $J$ is a judgement derived by the automatic analyis,
  then $\tr(J)$ is a quantitative Hoare triple derivable
  in the quantitative logic.
\end{theorem}
%
\noindent
The proof of this theorem is constructive and basically maps each rule
of the automatic analysis directly to its counterpart in the
quantitative logic.  The most tricky parts are the translations of the
rules for increments and decrements and the rule {\sc Q:Weak} for
weakening because they make essential use of the algebraic properties
of our linear potential functions.
\iffull{ For instance we prove the following lemma to show the soundness
of the translation of {\sc Q:Weak}.
